QPLIB: A Library of Quadratic Programming Instances

### 2017-01-23 QPLIB-2.2.tex
replace \bf comments by \todo

% -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -
\subsubsection{Classification}\label{ssec:taxonomy}

@@

When a $Q^i$ is positive semidefinite (PSD), i.e., the corresponding constraint/objective function is convex, general quadratic constraints are in fact equivalent to diagonal ones. In fact, every PSD matrix can be factorized as $Q^i = L^i (L^i)^\top$, e.g.~by the (incomplete) Cholesky factorization, $f^i(x) = x^\top Q^i x = \sum_{j \in \Mcn} z^2_jz$ where $z = x^\top L^i$. Hence, one could think that D** problems need not be distinguished from C** ones; however, this is true only for ``complicated'' constraints, but not for ``simple'' ones, because the above reformulation technique introduces linear constraints. Indeed, while C*L (and, a fortiori, C*\{C,Q\}) can always be brought to D*L (D*\{C,Q\}), using the same technique C*B becomes D*L, which is significantly different from D*B. In practice, a diagonal convex objective function under linear constraints is found in many applications ({\bf citations?}), so that D*L still makes sense to distinguish the case where the objective function is ``naturally'' separable from that where separability is artificially introduced, although this is in theory always possible.

@@

 \item \emph{Convex Continuous Quadratic Programs} CCC can be solved in polynomial time by Interior-Point techniques \cite{Wright97}; the simpler CCL can also be solved by means of ``simplex-like'' techniques, usually referred to as active-set methods \cite{Dost09}. Actually, a slightly larger class of problems can be solved with Interior-Point methods: those that can be represented by Second-Order Cone Programs. When written as QPs the corresponding $Q^i$ may not be positive semidefinite, but still the problems can be efficiently solved. Of course, these problems can still require considerable time, like LCL, when the size of the instance grows. In this sense, like in the linear case, a significant divide is from solvers that need all the data of QP to work, and those that are ``matrix-free'', i.e., only require the solution of simple operations (typically, matrix-vector products) with the data of the problem to work ({\bf citations?}). While in our library we have never exploited such a characteristic, which is not suitable to the use of standard modeling tools, this may be relevant for the solution of very-large-scale CIC.

@@

 \item \emph{Nonconvex Continuous Quadratic Programs} QCQ are instead in general $\mathcal{NP}$-hard, even if the constraints are very specific (QCB) and only one eigenvalue of $Q^0$ is negative \cite{Hemmecke2010}. They therefore require enumerative techniques like the spatial Branch\&Bound ({\bf citations?}), to be solved to optimality. Of course, local approaches are available that are able to efficiently provide saddle points (hopefully, local optima) of the CQC, but providing global guarantees about the quality of the obtained solutions is challenging. In our library we have specifically focussed on  \emph{exact} solution of the instances.

@@

 \item \emph{Convex Integer Quadratic Programs} CGC are in general $\mathcal{NP}$-hard, and therefore require enumerative techniques to be solved. However, convexity of the objective function and constraints implies that efficient techniques (see CCC) can be used at least to solve the continuous relaxation. The general view is that CGC are not, all other things being equal,  substantially more difficult than LGL to solve, especially if the objective function and/or the constraints have specific properties (e.g., DGL, CGL). Often integer variables are in fact binary ones, so several CCC models are C\{B,M\}C ones. In practice binary variables are considered to lead to somewhat easier problems than general integer ones (cf.~the cited result about hardness of unbounded integer quadratic programs), and several algorithmic techniques have been specifically developed for this special case. However, the general approaches for CBC are basically the same as for CGC, so there is seldom the need to distinguish between the two classes as far as solvability is concerned, although matters can be different regarding actual solution cost. Programs with only binary variables CBC can be easier than mixed-binary or integer ones C\{M,I\}C because some techniques are specifically known for the binary-only case, cf.~the next point ({\bf citations?}). Absence of continuous variables, even in the presence of integer ones CIC, can also lead to specific techniques ({\bf citations??}).  

@@

  \item \emph{Nonconvex Binary Quadratic Programs} QB\{B, N, L\} obviously are $\mathcal{NP}$-hard. However, the special nature of binary variables combined with quadratic forms allows for quite specific techniques to be developed, among which is the reformulation of the problem as a LBL. Also, many well-known combinatorial problems can be naturally reformulated as problems of this class, and therefore a considerable number of results have been obtained by exploiting specific properties of the set of constraints ( \cite{Rendl2008}, {\bf some citations? some specific problem to mention apart from Max-Cut?}).

@@

  \item \emph{Nonconvex Integer Quadratic Programs} QGQ is the most general, and therefore is the most difficul, class. Due to the lack of convexity even when integrality requirements are removed, solution methods must typically combine several algorithmic ideas, such as enumeration (distinguishing the role of integral variables from that of continuous ones involved into nonconvex terms) and techniques (e.g., outer approximation, SDP relaxation, \ldots) that allow to efficiently compute bounds. As in the convex case, QBQ, QMQ, and QIQ can benefit from more specific properties of the variables (\cite{Buchheim2013,Dong2016}, {\bf citations?}).

>>>>

% -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -  -
\subsubsection{Classification}\label{ssec:taxonomy}
\todo[inline]{subsubsection has some title as subsection}

@@

When a $Q^i$ is positive semidefinite (PSD), i.e., the corresponding constraint/objective function is convex, general quadratic constraints are in fact equivalent to diagonal ones. In fact, every PSD matrix can be factorized as $Q^i = L^i (L^i)^\top$, e.g.~by the (incomplete) Cholesky factorization, $f^i(x) = x^\top Q^i x = \sum_{j \in \Mcn} z^2_jz$ where $z = x^\top L^i$. Hence, one could think that D** problems need not be distinguished from C** ones; however, this is true only for ``complicated'' constraints, but not for ``simple'' ones, because the above reformulation technique introduces linear constraints. Indeed, while C*L (and, a fortiori, C*\{C,Q\}) can always be brought to D*L (D*\{C,Q\}), using the same technique C*B becomes D*L, which is significantly different from D*B. In practice, a diagonal convex objective function under linear constraints is found in many applications\todo{citations?}, so that D*L still makes sense to distinguish the case where the objective function is ``naturally'' separable from that where separability is artificially introduced, although this is in theory always possible.

@@

 \item \emph{Convex Continuous Quadratic Programs} CCC can be solved in polynomial time by Interior-Point techniques \cite{Wright97}; the simpler CCL can also be solved by means of ``simplex-like'' techniques, usually referred to as active-set methods \cite{Dost09}. Actually, a slightly larger class of problems can be solved with Interior-Point methods: those that can be represented by Second-Order Cone Programs. When written as QPs the corresponding $Q^i$ may not be positive semidefinite, but still the problems can be efficiently solved. Of course, these problems can still require considerable time, like LCL, when the size of the instance grows. In this sense, like in the linear case, a significant divide is from solvers that need all the data of QP to work, and those that are ``matrix-free'', i.e., only require the solution of simple operations (typically, matrix-vector products) with the data of the problem to work\todo{citations?}. While in our library we have never exploited such a characteristic, which is not suitable to the use of standard modeling tools, this may be relevant for the solution of very-large-scale CIC.

@@

 \item \emph{Nonconvex Continuous Quadratic Programs} QCQ are instead in general $\mathcal{NP}$-hard, even if the constraints are very specific (QCB) and only one eigenvalue of $Q^0$ is negative \cite{Hemmecke2010}. They therefore require enumerative techniques like the spatial Branch\&Bound\todo{citations?}, to be solved to optimality. Of course, local approaches are available that are able to efficiently provide saddle points (hopefully, local optima) of the CQC, but providing global guarantees about the quality of the obtained solutions is challenging. In our library we have specifically focussed on  \emph{exact} solution of the instances.

@@

 \item \emph{Convex Integer Quadratic Programs} CGC are in general $\mathcal{NP}$-hard, and therefore require enumerative techniques to be solved. However, convexity of the objective function and constraints implies that efficient techniques (see CCC) can be used at least to solve the continuous relaxation. The general view is that CGC are not, all other things being equal,  substantially more difficult than LGL to solve, especially if the objective function and/or the constraints have specific properties (e.g., DGL, CGL). Often integer variables are in fact binary ones, so several CCC models are C\{B,M\}C ones. In practice binary variables are considered to lead to somewhat easier problems than general integer ones (cf.~the cited result about hardness of unbounded integer quadratic programs), and several algorithmic techniques have been specifically developed for this special case. However, the general approaches for CBC are basically the same as for CGC, so there is seldom the need to distinguish between the two classes as far as solvability is concerned, although matters can be different regarding actual solution cost. Programs with only binary variables CBC can be easier than mixed-binary or integer ones C\{M,I\}C because some techniques are specifically known for the binary-only case, cf.~the next point\todo{citations?}. Absence of continuous variables, even in the presence of integer ones CIC, can also lead to specific techniques\todo{citations??}.

@@

  \item \emph{Nonconvex Binary Quadratic Programs} QB\{B, N, L\} obviously are $\mathcal{NP}$-hard. However, the special nature of binary variables combined with quadratic forms allows for quite specific techniques to be developed, among which is the reformulation of the problem as a LBL. Also, many well-known combinatorial problems can be naturally reformulated as problems of this class, and therefore a considerable number of results have been obtained by exploiting specific properties of the set of constraints \cite{Rendl2008}\todo{some citations? some specific problem to mention apart from Max-Cut?}.

@@

  \item \emph{Nonconvex Integer Quadratic Programs} QGQ is the most general, and therefore is the most difficul, class. Due to the lack of convexity even when integrality requirements are removed, solution methods must typically combine several algorithmic ideas, such as enumeration (distinguishing the role of integral variables from that of continuous ones involved into nonconvex terms) and techniques (e.g., outer approximation, SDP relaxation, \ldots) that allow to efficiently compute bounds. As in the convex case, QBQ, QMQ, and QIQ can benefit from more specific properties of the variables \cite{Buchheim2013,Dong2016}\todo{citations?}.

### 2017-01-23 QPLIB-3.tex
replace \bf comments by \todo

During the development of the library, a discussion ensued about what
the expected gols that we wanted to achieve. The following four goals
where finally identified:
%
\begin{enumerate}
 \item represent as much as possible all the different categories of QP
       problems;
 %
 \item gather ``challenging'' instances, i.e., ones which can not be easily
       solved by  state-of-the-art solvers;
 %
 \item include for each of the categories a set of well-diversified
       instances;
 %
 \item obtain a set of instances which is neither too small, so as to
       preserve statistical relevance, nor too large so as to being
       computationally manageable.
\end{enumerate}
%
To achieve the aforementioned goals, we performed the following two
filters, applied in cascade.
%
\begin{itemize}
 \item \emph{First Instances Filter.}\\
       The first filter was designed to drastically reduce the number of
       instances by eliminating the ``easy'' ones. An empirical measure
       for the hardness of an instance is the CPU time needed by a
       complete solver (cf.~\S~\ref{sec:algo}) to solve it to
       global optimality. Accordingly, for each of the gathered instance we
       ran the complete solvers in {\tt GAMS} (cf.~Table \ref{t:solvers})
       capable of solving it, whose number depends on the category of the
       instance under consideration. We then filtered according to a relative
       measure of computational difficulty, i.e., we discarded all instances
       that are solved by at least 30\% of the complete solvers within a time
       limit of 30 seconds.
 %
 \item \emph{Second Instances Filter.}\\
       The goal of the second filter was to eliminate ``similar'' instances.
       We carefully analysed the instances one by one, and we clustered them
       according to their features; for each cluster we kept only a few
       representatives. {\bf Comment: we may want to be a tad more specific
       about what ``features'' exactly meant there}. Finally, in order to only
       keep computationally challenging instances we ran a general purpose solver
       able to tackle all QP categories with a time limit of 120 seconds; all the
       instances which have been solved to proven optimality within this time limit
       were discarded. {\bf Comment: this might actually be best characterized as
       a third filter.}
\end{itemize}
%
In Table~\ref{tab:filters} we summarize the two filter steps, which
allowed us to identify the final set of $251$ discrete instances and
$116$ continuous instances.

>>>>

During the development of the library, a discussion ensued about what
the expected gols that we wanted to achieve. The following four goals
where finally identified:
%
\begin{enumerate}
 \item represent as much as possible all the different categories of QP
       problems;
 %
 \item gather ``challenging'' instances, i.e., ones which can not be easily
       solved by  state-of-the-art solvers;
 %
 \item include for each of the categories a set of well-diversified
       instances;
 %
 \item obtain a set of instances which is neither too small, so as to
       preserve statistical relevance, nor too large so as to being
       computationally manageable.
\end{enumerate}
%
To achieve the aforementioned goals, we performed the following two
filters, applied in cascade.
%
\begin{itemize}
 \item \emph{First Instances Filter.}\\
       The first filter was designed to drastically reduce the number of
       instances by eliminating the ``easy'' ones. An empirical measure
       for the hardness of an instance is the CPU time needed by a
       complete solver (cf.~\S~\ref{sec:algo}) to solve it to
       global optimality. Accordingly, for each of the gathered instance we
       ran the complete solvers in {\tt GAMS} (cf.~Table \ref{t:solvers})
       capable of solving it, whose number depends on the category of the
       instance under consideration. We then filtered according to a relative
       measure of computational difficulty, i.e., we discarded all instances
       that are solved by at least 30\% of the complete solvers within a time
       limit of 30 seconds.
 %
 \item \emph{Second Instances Filter.}\\
       The goal of the second filter was to eliminate ``similar'' instances.
       We carefully analysed the instances one by one, and we clustered them
       according to their features; for each cluster we kept only a few
       representatives\todo{we may want to be a tad more specific
       about what ``features'' exactly meant there}. Finally, in order to only
       keep computationally challenging instances we ran a general purpose solver
       able to tackle all QP categories with a time limit of 120 seconds; all the
       instances which have been solved to proven optimality within this time limit
       were discarded\todo{this might actually be best characterized as
       a third filter}.
\end{itemize}
%
In Table~\ref{tab:filters} we summarize the two filter steps, which
allowed us to identify the final set of $251$ discrete instances and
$116$ continuous instances.

### 2017-01-23 methods.tex
replace \bf comments by \todo

In general, the applicability of meta-heuristics to a given problem depends 
on whether the local search they utilize
%on what whether the local search they rest on 
addresses that problem or not. 
{\bf Antonio: the concept of ``addresses that problem'' is not very clear to me, this sentence may benefit from rephrasing (or the paragraph from deleting if it's not deemed to be utterly necessary).} Depending on the local search employed, Multi-start methods can address MINLP of the most general class.

>>>>

In general, the applicability of meta-heuristics to a given problem depends 
on whether the local search they utilize
%on what whether the local search they rest on 
addresses that problem or not. 
\todo[inline]{Antonio: the concept of ``addresses that problem'' is not very clear to me, this sentence may benefit from rephrasing (or the paragraph from deleting if it's not deemed to be utterly necessary).} Depending on the local search employed, Multi-start methods can address MINLP of the most general class.
